---
title: "Untitled"
output: html_document
date: "2024-03-06"
---

This report analyses the prevalence of obesity in Scotland, utilizing data from the Scottish Health Surveys conducted between 2008 and 2012. The data encompass socio-economic and lifestyle factors including age, sex, highest educational qualification, daily vegetable and fruit intake, and obesity classification. The main objectives of this analysis are to determine whether the prevalence of obesity has changed over the survey years and to identify any differences in obesity by age, gender, socio-economic status, and lifestyle factors.

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(ggplot2)

# Load the data
data <- read_csv("DAProject1.csv")
```


```{r}
data$Sex <- as.factor(data$Sex)
data$Education <- as.factor(data$Education)
data$Veg <- as.factor(ifelse(data$Veg == "Yes", 1, 0))
data$Fruit <- as.factor(ifelse(data$Fruit == "Yes", 1, 0))
data$Obese <- as.factor(ifelse(data$Obese == "Yes", 1, 0))
```

```{r}
# Assuming 'Obese' is a factor with levels 'Yes' and 'No'
# Calculate the annual prevalence of obesity
annual_prevalence <- data %>%
  group_by(Year) %>%
  summarise(ObesityPrevalence = mean(Obese == "Yes", na.rm = TRUE))

# Plot the trend over years
ggplot(annual_prevalence, aes(x = Year, y = ObesityPrevalence)) +
  geom_line(aes(group=1), colour="blue") +
  geom_point(colour="red") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Obesity Prevalence Over the Years", x = "Year", y = "Prevalence (%)")

```




```{r}
# Logistic regression with Age as a predictor
model_age <- glm(Obese ~ Age, family = binomial(link = "logit"), data = data)
summary(model_age)

# Logistic regression with Sex as a predictor
model_sex <- glm(Obese ~ Sex, family = binomial(link = "logit"), data = data)
summary(model_sex)

```

```{r}
# Multiple logistic regression
model_full <- glm(Obese ~ Age + Sex + Education + Veg + Fruit, family = binomial(link = "logit"), data = data)
summary(model_full)

```

```{r}
library(ggplot2)
data$predicted_prob <- predict(model_age, type = "response")
ggplot(data, aes(x = Age, y = predicted_prob)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(title = "Predicted Probability of Obesity by Age", x = "Age", y = "Predicted Probability")

```

```{r}
obesity_prevalence <- data %>%
  group_by(Year) %>%
  summarise(ObesityPrevalence = mean(Obese == "Yes", na.rm = TRUE))

# Line chart for obesity prevalence over the years
ggplot(obesity_prevalence, aes(x = Year, y = ObesityPrevalence)) +
  geom_line() + 
  geom_point() + # Adding points to the line chart for each year
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Trend of Obesity Prevalence Over the Years", x = "Year", y = "Obesity Prevalence (%)")

# Obesity by age and gender
ggplot(data, aes(x = Age, fill = Obese)) +
  geom_histogram(position = "fill", bins = 30) +
  facet_wrap(~Sex) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Obesity by Age and Gender", x = "Age", y = "Percentage")

# Checking for missing values
sum(is.na(data))

```
```{r}
# Exploratory Data Analysis (EDA)

## Load necessary libraries
library(tidyverse)

## Load the data
survey_data <- read.csv("DAProject1.csv")

## Preview the data
head(survey_data)

## Summary statistics
summary(survey_data)

## Visualize the distribution of obesity over the years
ggplot(survey_data, aes(x=Year, by=Obese, fill=Obese)) +
  geom_bar(position="fill") +
  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Obesity Prevalence Over the Years", y="Percentage", x="Year")

## Obesity prevalence by gender
ggplot(survey_data, aes(x=Sex, by=Obese, fill=Obese)) +
  geom_bar(position="fill") +
  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Obesity Prevalence by Gender", y="Percentage", x="Sex")

```
```{r}
# Load necessary libraries
library(tidyverse)

# Assuming survey_data is already loaded and contains the dataset

## Filter data for individuals not obese and obese separately
not_obese_data <- survey_data %>% filter(Obese == "No")
obese_data <- survey_data %>% filter(Obese == "Yes")

## Visualize the impact of vegetable intake on those not classified as obese
ggplot(not_obese_data, aes(x=Veg)) +
  geom_bar() +
  labs(title="Vegetable Intake among Individuals Not Classified as Obese", y="Count", x="Vegetable Intake")

## Visualize the impact of vegetable intake on those classified as obese
ggplot(obese_data, aes(x=Veg)) +
  geom_bar() +
  labs(title="Vegetable Intake among Individuals Classified as Obese", y="Count", x="Vegetable Intake")

```

```{r}
# Formal Analysis

## Obesity prevalence by education level
ggplot(survey_data, aes(x=Education, fill=Obese)) +
  geom_bar(position="fill") +
  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Obesity Prevalence by Education Level", y="Percentage", x="Education") +
  theme(axis.text.x = element_text(angle=45, hjust=1))

## Analyze the impact of lifestyle factors on obesity
ggplot(survey_data, aes(x=Veg, fill=Obese)) +
  geom_bar(position="fill") +
  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Impact of Vegetable Intake on Obesity Prevalence", y="Percentage", x="Vegetable Intake")

ggplot(survey_data, aes(x=Fruit, fill=Obese)) +
  geom_bar(position="fill") +
  scale_y_continuous(labels=scales::percent_format()) +
  labs(title="Impact of Fruit Intake on Obesity Prevalence", y="Percentage", x="Fruit Intake")

```

